Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-10-22 Methods-and-Results.Rmd
forest and overall ols added

>>>>

\section{Embeddedness and Stability across Studies}

\begin{center}
    \noindent\rule{8cm}{0.4pt}\\
    \hl{\textit{New Section:}}\\
    \vspace{-0.25cm}
    \noindent\rule{8cm}{0.4pt}
\end{center}

Introduction sentence

\hl{maybe move this first paragraph to the end?}
\subsection{Stability}
We begin by assessing the stability of our main analyses across the three studies. To do so, we plot the effect sizes of each parameter of interest in a forest plot. Forest plots are a popular form of evidence synthesis and are commonly found within meta-analytic efforts. By plotting a standardized effect size and confidence interval of key parameters for each study, forest plots offer a visual method of identifying broader patterns across studies. This affords users a chance to spot stable as well as unreliable effects more easily. While not the focus of our efforts here, forest plots are commonly accompanied with a meta-analytic effect size that summarizes the effect size across studies.

We assess forest plots of the basic contact hypothesis test (Figure \ref{fig:ContactHypothesis}), the core need fulfillment proposal (Figure \ref{fig:AllportNeedFulfillment}), as well as the robustness analyses (Figure \ref{fig:Robustness}). For the basic contact hypothesis test we see that outgroup contact has a strong and consistent effect on outgroup attitudes while interactions with non-outgroup members consistently have no meaningful effect on outgroup attitudes. While this would be expected from the general intergroup contact literature, this is not a trivial finding. We are among the first to assess the contact hypothesis using real-life data from intensive longitudinal data and these close to event assessments are comparable but not equivalent to the assessments found in cross-sectional data (also see below). When looking at Figure \ref{fig:AllportNeedFulfillment}, we see that the fulfillment of core needs during intergroup contacts is consistently a meaningful predictor of outgroup attitudes and interaction quality perceptions (see Figure \ref{fig:AllportNeedFulfillment} A and B). We also see that the effect of core need fulfillment on outgroup attitudes is strongly reduced when modeled together with interaction quality perceptions, supporting our assertion that interaction quality offers a psychological mechanism for the direct effect. Additionally, the core need fulfillment is also consistently outgroup contact specific (Figure \ref{fig:Robustness} A) and remains a meaningful predictor of outgroup attitudes even when controlling for traditional psychological need fulfillments (Figure \ref{fig:Robustness} B).

\subsection{General contact hypothesis}
```{r join data sets for analysis}
# Join aggregated data
avS1 <- dtWorkerSupp$workerAvFreQual %>%
  select(
    -ExternalReference
  ) %>%
  mutate(
    PID = c(1:nrow(.)),
    study = "S1"
  ) %>%
  select(PID, everything())
avS2 <- dtStudentSupp$studentContactFreq %>%
  mutate(
    study = "S2"
  )
avS3 <- dtMedicalSupp$medicalContactFreq %>%
  mutate(
    study = "S3"
  )

avAll <- bind_rows(avS1, avS2, avS3) %>%
  mutate(
    ID = c(1:nrow(.)),
    studyNum1 = ifelse(study == "S1", 1, 0),
    studyNum3 = ifelse(study == "S3", 1, 0)
  )
avAll$study2 <- relevel(as.factor(avAll$study),"S2")
```
```{r OLS regression across the three studies}
lmAllAttFreqQualX <-
  lm(
    AvAttitude ~ SumContactNL_c * AvQuality_c + SumContactNL_c * study2 + AvQuality_c * study2,
    data = avAll
  )

summ(
  lmAllAttFreqQualX,
  confint = TRUE,
  digits = 3
)

# prepare F stat for manuscript
fit <- summary(lmAllAttFreqQualX)
lmAllFstatLatex <- paste(
  "\\textit{F}(",
  fit$fstatistic["numdf"],
  ", ",
  fit$fstatistic["dendf"],
  ") = ",
  format(round(fit$fstatistic["value"], 2), nsmall = 2),
  ", \\textit{p} = ",
  format(round(
    pf(
      fit$fstatistic["value"],
      fit$fstatistic["numdf"],
      fit$fstatistic["dendf"],
      lower.tail = FALSE
    ),
    3
  ), nsmall = 3),
  ", \\textit{R\\textsuperscript{2}} = ",
  format(round(fit$r.squared, 2), nsmall = 2),
  sep = ""
)
```

During the main analyses, we have thus far shown that participants held more positive outgroup attitudes following intergroup contacts and that a perceived interaction quality was associated with more positive outgroup attitudes following an intergroup contact. We have, however, not brought the two elements of the contact hypothesis together in a single analysis. This is mainly due to the difficulty that interaction quality ratings are only possible when participants had an interaction --- which means that the two factors cannot be assessed in a multilevel analysis that models individual measurement occasions. This has not been an issue in the past literature because experimental and cross-sectional data collections have only considered cases in which intergroup contacts were present. Outside of the lab, the most common method has been to ask participants to recall how many interactions they had over the past month, how positive these interactions were, and how positive they feel towards the outgroup in general. The effects of intergroup contacts, interaction quality, and outgroup attitudes are thus aggregated by the participant in a cognitive recall aggregation. Intensive longitudinal data also allows for aggregation of all three variables, with the differences that the measurements are aggregated from individual reports that are close to the actual experience. However, this aggregation process summarizes the dense multitude of measurement occasions as single summary statistics per participants. This by extension means that we face statistical power restrictions on the person level and have this decided to only present this analysis based on a combined data set of all three studies.

We thus run a linear regression model where average outgroup attitudes are predicted by the number of interactions and the average interaction quality ratings of all participants across the three studies. We do so while controlling for the possible effects of study-specific differences. To include the study-membership as a control variable, we use the student sample (Study 2) as the reference group because it is both the largest and the most homogeneous study. Looking at the overall model, we find that the model predicts a statistically significant amount of variance in average outgroup attitudes (`r lmAllFstatLatex`). Looking at the individual effects, we find that only the number of outgroup interactions is associated with average outgroup attitudes (`r LmCoeffLatex(lmMdl = lmAllAttFreqQualX, varName = "SumContactNL_c", ci = TRUE)`). Neither the average interaction quality perceptions nor an interaction of amount and quality of contact are meaningful or significant predictors of average outgroup attitudes (both \textit{b} <0.22 and \textit{p} > .201). In short, participants who on average had more positive interactions were not statistically more likely to show higher average outgroup attitudes. This result is surprising because it contradicts the tenet of the intergroup contact hypothesis that intergroup contact depends on positive contact. Additionally, the aggregated results contradict the results of our own multi-level regressions, where we have shown that the positive effect of individual outgroup interactions dependent on interaction quality perceptions.

\subsection{Contact key need content}
topic model

### 2022-10-22 Methods-and-Results.tex
forest and overall ols added

>>>>

\section{Embeddedness and Stability across Studies}

\begin{center}
    \noindent\rule{8cm}{0.4pt}\\
    \hl{\textit{New Section:}}\\
    \vspace{-0.25cm}
    \noindent\rule{8cm}{0.4pt}
\end{center}

Introduction sentence

\hl{maybe move this first paragraph to the end?}
\subsection{Stability}

We begin by assessing the stability of our main analyses across the
three studies. To do so, we plot the effect sizes of each parameter of
interest in a forest plot. Forest plots are a popular form of evidence
synthesis and are commonly found within meta-analytic efforts. By
plotting a standardized effect size and confidence interval of key
parameters for each study, forest plots offer a visual method of
identifying broader patterns across studies. This affords users a chance
to spot stable as well as unreliable effects more easily. While not the
focus of our efforts here, forest plots are commonly accompanied with a
meta-analytic effect size that summarizes the effect size across
studies.

We assess forest plots of the basic contact hypothesis test (Figure
\ref{fig:ContactHypothesis}), the core need fulfillment proposal (Figure
\ref{fig:AllportNeedFulfillment}), as well as the robustness analyses
(Figure \ref{fig:Robustness}). For the basic contact hypothesis test we
see that outgroup contact has a strong and consistent effect on outgroup
attitudes while interactions with non-outgroup members consistently have
no meaningful effect on outgroup attitudes. While this would be expected
from the general intergroup contact literature, this is not a trivial
finding. We are among the first to assess the contact hypothesis using
real-life data from intensive longitudinal data and these close to event
assessments are comparable but not equivalent to the assessments found
in cross-sectional data (also see below). When looking at Figure
\ref{fig:AllportNeedFulfillment}, we see that the fulfillment of core
needs during intergroup contacts is consistently a meaningful predictor
of outgroup attitudes and interaction quality perceptions (see Figure
\ref{fig:AllportNeedFulfillment} A and B). We also see that the effect
of core need fulfillment on outgroup attitudes is strongly reduced when
modeled together with interaction quality perceptions, supporting our
assertion that interaction quality offers a psychological mechanism for
the direct effect. Additionally, the core need fulfillment is also
consistently outgroup contact specific (Figure \ref{fig:Robustness} A)
and remains a meaningful predictor of outgroup attitudes even when
controlling for traditional psychological need fulfillments (Figure
\ref{fig:Robustness} B).

\subsection{General contact hypothesis}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Join aggregated data}
\NormalTok{avS1 }\OtherTok{\textless{}{-}}\NormalTok{ dtWorkerSupp}\SpecialCharTok{$}\NormalTok{workerAvFreQual }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(}
    \SpecialCharTok{{-}}\NormalTok{ExternalReference}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{PID =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{nrow}\NormalTok{(.)),}
    \AttributeTok{study =} \StringTok{"S1"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(PID, }\FunctionTok{everything}\NormalTok{())}
\NormalTok{avS2 }\OtherTok{\textless{}{-}}\NormalTok{ dtStudentSupp}\SpecialCharTok{$}\NormalTok{studentContactFreq }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{study =} \StringTok{"S2"}
\NormalTok{  )}
\NormalTok{avS3 }\OtherTok{\textless{}{-}}\NormalTok{ dtMedicalSupp}\SpecialCharTok{$}\NormalTok{medicalContactFreq }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{study =} \StringTok{"S3"}
\NormalTok{  )}

\NormalTok{avAll }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(avS1, avS2, avS3) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{ID =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{nrow}\NormalTok{(.)),}
    \AttributeTok{studyNum1 =} \FunctionTok{ifelse}\NormalTok{(study }\SpecialCharTok{==} \StringTok{"S1"}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
    \AttributeTok{studyNum3 =} \FunctionTok{ifelse}\NormalTok{(study }\SpecialCharTok{==} \StringTok{"S3"}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{)}
\NormalTok{  )}
\NormalTok{avAll}\SpecialCharTok{$}\NormalTok{study2 }\OtherTok{\textless{}{-}} \FunctionTok{relevel}\NormalTok{(}\FunctionTok{as.factor}\NormalTok{(avAll}\SpecialCharTok{$}\NormalTok{study),}\StringTok{"S2"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{lmAllAttFreqQualX }\OtherTok{\textless{}{-}}
  \FunctionTok{lm}\NormalTok{(}
\NormalTok{    AvAttitude }\SpecialCharTok{\textasciitilde{}}\NormalTok{ SumContactNL\_c }\SpecialCharTok{*}\NormalTok{ AvQuality\_c }\SpecialCharTok{+}\NormalTok{ SumContactNL\_c }\SpecialCharTok{*}\NormalTok{ study2 }\SpecialCharTok{+}\NormalTok{ AvQuality\_c }\SpecialCharTok{*}\NormalTok{ study2,}
    \AttributeTok{data =}\NormalTok{ avAll}
\NormalTok{  )}

\FunctionTok{summ}\NormalTok{(}
\NormalTok{  lmAllAttFreqQualX,}
  \AttributeTok{confint =} \ConstantTok{TRUE}\NormalTok{,}
  \AttributeTok{digits =} \DecValTok{3}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]
\centering
\begin{tabular}{lr}
\toprule
\cellcolor{gray!6}{Observations} & \cellcolor{gray!6}{205 (2 missing obs. deleted)}\\
Dependent variable & AvAttitude\\
\cellcolor{gray!6}{Type} & \cellcolor{gray!6}{OLS linear regression}\\
\bottomrule
\end{tabular}
\end{table} \begin{table}[!h]
\centering
\begin{tabular}{lr}
\toprule
\cellcolor{gray!6}{F(9,195)} & \cellcolor{gray!6}{2.643}\\
R² & 0.109\\
\cellcolor{gray!6}{Adj. R²} & \cellcolor{gray!6}{0.068}\\
\bottomrule
\end{tabular}
\end{table} \begin{table}[!h]
\centering
\begin{threeparttable}
\begin{tabular}{lrrrrr}
\toprule
  & Est. & 2.5\% & 97.5\% & t val. & p\\
\midrule
\cellcolor{gray!6}{(Intercept)} & \cellcolor{gray!6}{67.297} & \cellcolor{gray!6}{64.274} & \cellcolor{gray!6}{70.319} & \cellcolor{gray!6}{43.908} & \cellcolor{gray!6}{0.000}\\
SumContactNL\_c & 0.736 & 0.328 & 1.144 & 3.558 & 0.000\\
\cellcolor{gray!6}{AvQuality\_c} & \cellcolor{gray!6}{0.222} & \cellcolor{gray!6}{-0.119} & \cellcolor{gray!6}{0.563} & \cellcolor{gray!6}{1.283} & \cellcolor{gray!6}{0.201}\\
study2S1 & 3.062 & -4.600 & 10.724 & 0.788 & 0.432\\
\cellcolor{gray!6}{study2S3} & \cellcolor{gray!6}{-2.510} & \cellcolor{gray!6}{-7.348} & \cellcolor{gray!6}{2.328} & \cellcolor{gray!6}{-1.023} & \cellcolor{gray!6}{0.307}\\
\addlinespace
SumContactNL\_c:AvQuality\_c & 0.002 & -0.020 & 0.024 & 0.198 & 0.843\\
\cellcolor{gray!6}{SumContactNL\_c:study2S1} & \cellcolor{gray!6}{-0.684} & \cellcolor{gray!6}{-1.441} & \cellcolor{gray!6}{0.073} & \cellcolor{gray!6}{-1.782} & \cellcolor{gray!6}{0.076}\\
SumContactNL\_c:study2S3 & -0.595 & -1.076 & -0.113 & -2.437 & 0.016\\
\cellcolor{gray!6}{AvQuality\_c:study2S1} & \cellcolor{gray!6}{0.518} & \cellcolor{gray!6}{-0.313} & \cellcolor{gray!6}{1.349} & \cellcolor{gray!6}{1.229} & \cellcolor{gray!6}{0.220}\\
AvQuality\_c:study2S3 & 0.133 & -0.367 & 0.634 & 0.526 & 0.600\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item Standard errors: OLS
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# prepare F stat for manuscript}
\NormalTok{fit }\OtherTok{\textless{}{-}} \FunctionTok{summary}\NormalTok{(lmAllAttFreqQualX)}
\NormalTok{lmAllFstatLatex }\OtherTok{\textless{}{-}} \FunctionTok{paste}\NormalTok{(}
  \StringTok{"}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{textit\{F\}("}\NormalTok{,}
\NormalTok{  fit}\SpecialCharTok{$}\NormalTok{fstatistic[}\StringTok{"numdf"}\NormalTok{],}
  \StringTok{", "}\NormalTok{,}
\NormalTok{  fit}\SpecialCharTok{$}\NormalTok{fstatistic[}\StringTok{"dendf"}\NormalTok{],}
  \StringTok{") = "}\NormalTok{,}
  \FunctionTok{format}\NormalTok{(}\FunctionTok{round}\NormalTok{(fit}\SpecialCharTok{$}\NormalTok{fstatistic[}\StringTok{"value"}\NormalTok{], }\DecValTok{2}\NormalTok{), }\AttributeTok{nsmall =} \DecValTok{2}\NormalTok{),}
  \StringTok{", }\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{textit\{p\} = "}\NormalTok{,}
  \FunctionTok{format}\NormalTok{(}\FunctionTok{round}\NormalTok{(}
    \FunctionTok{pf}\NormalTok{(}
\NormalTok{      fit}\SpecialCharTok{$}\NormalTok{fstatistic[}\StringTok{"value"}\NormalTok{],}
\NormalTok{      fit}\SpecialCharTok{$}\NormalTok{fstatistic[}\StringTok{"numdf"}\NormalTok{],}
\NormalTok{      fit}\SpecialCharTok{$}\NormalTok{fstatistic[}\StringTok{"dendf"}\NormalTok{],}
      \AttributeTok{lower.tail =} \ConstantTok{FALSE}
\NormalTok{    ),}
    \DecValTok{3}
\NormalTok{  ), }\AttributeTok{nsmall =} \DecValTok{3}\NormalTok{),}
  \StringTok{", }\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{textit\{R}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{textsuperscript\{2\}\} = "}\NormalTok{,}
  \FunctionTok{format}\NormalTok{(}\FunctionTok{round}\NormalTok{(fit}\SpecialCharTok{$}\NormalTok{r.squared, }\DecValTok{2}\NormalTok{), }\AttributeTok{nsmall =} \DecValTok{2}\NormalTok{),}
  \AttributeTok{sep =} \StringTok{""}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

During the main analyses, we have thus far shown that participants held
more positive outgroup attitudes following intergroup contacts and that
a perceived interaction quality was associated with more positive
outgroup attitudes following an intergroup contact. We have, however,
not brought the two elements of the contact hypothesis together in a
single analysis. This is mainly due to the difficulty that interaction
quality ratings are only possible when participants had an interaction
--- which means that the two factors cannot be assessed in a multilevel
analysis that models individual measurement occasions. This has not been
an issue in the past literature because experimental and cross-sectional
data collections have only considered cases in which intergroup contacts
were present. Outside of the lab, the most common method has been to ask
participants to recall how many interactions they had over the past
month, how positive these interactions were, and how positive they feel
towards the outgroup in general. The effects of intergroup contacts,
interaction quality, and outgroup attitudes are thus aggregated by the
participant in a cognitive recall aggregation. Intensive longitudinal
data also allows for aggregation of all three variables, with the
differences that the measurements are aggregated from individual reports
that are close to the actual experience. However, this aggregation
process summarizes the dense multitude of measurement occasions as
single summary statistics per participants. This by extension means that
we face statistical power restrictions on the person level and have this
decided to only present this analysis based on a combined data set of
all three studies.

We thus run a linear regression model where average outgroup attitudes
are predicted by the number of interactions and the average interaction
quality ratings of all participants across the three studies. We do so
while controlling for the possible effects of study-specific
differences. To include the study-membership as a control variable, we
use the student sample (Study 2) as the reference group because it is
both the largest and the most homogeneous study. Looking at the overall
model, we find that the model predicts a statistically significant
amount of variance in average outgroup attitudes (\textit{F}(9, 195) =
2.64, \textit{p} = 0.007, \textit{R\textsuperscript{2}} = 0.11). Looking
at the individual effects, we find that only the number of outgroup
interactions is associated with average outgroup attitudes (\textit{b} =
0.74, \textit{t}(195) = 3.56, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.33, 1.14{]}). Neither the average interaction
quality perceptions nor an interaction of amount and quality of contact
are meaningful or significant predictors of average outgroup attitudes
(both \textit{b} \textless0.22 and \textit{p} \textgreater{} .201). In
short, participants who on average had more positive interactions were
not statistically more likely to show higher average outgroup attitudes.
This result is surprising because it contradicts the tenet of the
intergroup contact hypothesis that intergroup contact depends on
positive contact. Additionally, the aggregated results contradict the
results of our own multi-level regressions, where we have shown that the
positive effect of individual outgroup interactions dependent on
interaction quality perceptions.

\subsection{Contact key need content}

topic model
